---
title: "Covid19 vaccine study Fall2021"
author: "Put your name here!"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="fall21.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- TheData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>% mutate(across(where(is.character),as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ =="gay")
DrugUser = filter(FinalData, DrugUser =="yes")
```


# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males


$H_0$ Males who take Vaccine will show its effective against contracting COVID-19

$H_A$ Males taking the Vaccine will show no difference in contracting COVID -19


### Barplot



```{r}
TheData <- guy %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())
basicC <- ggplot(TheData,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot
basicCC <- ggplot(TheData,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents  guy(males) who took the vaccine in blue  and the placebo vaccine in red we are trying to determine if eitheir taking the vaccine is effective in preventing COVID -19 or do you contract it. About 58% of the males who took the vaccine did get COVID -19 unlike those getting the placebo effect only about 42% ended up contracting COVID -19.The vaccine appears to be ineffective in protection against covid-19.




### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```


The numerical results show specific percentages of guys(men) that got COVID -19 after taking the placebo and taking the vaccine. Among the guys, 58% got COVID-19 after taking the vaccine, while the approximately 42% of guys that were given the placebo contracted covid-19.Showing a 16 percent difference .The vaccine is clearly not effective in this group of people .



### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
guyfish <- fisher.test(table1)
guyfish
```


The results of the fisher exact test shows  the odd ratio being 0.68 while the p vaule is 6e-04 a small number that the results are highly unlikely to occurr.


```{r}
effect <- guyfish$p.value/(guyfish$p.value +1)
effectivenessguy = guyfish$estimate/(guyfish$estimate +1)*100
print(effectivenessguy)
```

## Effectiveness for guy is `r round(effectivenessguy,0)`


```{r}
low=(guyfish$conf.int[1]/(guyfish$conf.int[1] +1))*100
hi=(guyfish$conf.int[2]/(guyfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the COIVD-19 vaccine for guys(males) which is 40.4. If the effectiveness is above 50% that means that the vaccine is effective however based on the 40.4 the vaccine is not effective in this gorup. The odds ratio is approximately how effective the vaccine is  40.Based on the interval 35.09 - 45.93, we can conclude that the vaccine is ineffective in protecting guys from getting covid-19 and the placebo is more effective in this group.







# GALS

$H_0$ GALS who take Vaccine will show its effective against contracting COVID-19

$H_A$ GALS taking the Vaccine will show no difference in contracting COVID -19





```{r}
TheData <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
# the group_by followed by summarize(count=n())
basicC <- ggplot(TheData,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(TheData,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```



### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```

### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
galfish <- fisher.test(table1)
galfish
```


```{r}
effect <- galfish$p.value/(galfish$p.value +1)
effectivenessgal = galfish$estimate/(galfish$estimate +1)*100
print(effectivenessgal)
```

## Effectiveness for GAl is `r round(effectivenessgal,0)`


```{r}
low=(galfish$conf.int[1]/(galfish$conf.int[1] +1))*100
hi=(galfish$conf.int[2]/(galfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`




# Drug Users

$H_0$ DrugUsers who take Vaccine will show its effective against contracting COVID-19

$H_A$ DrugUsers taking the Vaccine will show no difference in contracting COVID -19



```{r}
TheData <- DrugUser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
# the group_by followed by summarize(count=n())
basicC <- ggplot(TheData,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(TheData,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
DrugUserfish <- fisher.test(table1)
DrugUserfish
```



```{r}
effect <- DrugUserfish$p.value/(DrugUserfish$p.value +1)
effectivenessDrugUser = DrugUserfish$estimate/(DrugUserfish$estimate +1)*100
print(effectivenessDrugUser)
```

## Effectiveness for DrugUser is `r round(effectivenessguy,0)`


```{r}
low=(DrugUserfish$conf.int[1]/(DrugUserfish$conf.int[1] +1))*100
hi=(DrugUserfish$conf.int[2]/(DrugUserfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`




# LGBTQ

$H_0$ LGBTQ (Males) who take Vaccine will show its effective against contracting COVID-19

$H_A$ LGBTQ (Males) taking the Vaccine will show no difference in contracting COVID -19

### Barplot



```{r}
TheData <-gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
# the group_by followed by summarize(count=n())
basicC <- ggplot(TheData,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(TheData,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```



### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```



### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
gayfish <- fisher.test(table1)
gayfish
```

```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```

## Effectiveness for GAY is `r round(effectivenessgay,0)`


```{r}
low=(gayfish$conf.int[1]/(gayfish$conf.int[1] +1))*100
hi=(gayfish$conf.int[2]/(gayfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`




# Conclusions and recommendations

